---
title: "Homework2 Sta 601"
author: "Lianghui Li"
date: "9/10/2017"
output:
  pdf_document: 
    latex_engine: xelatex
  html_document: default
---

```{r setup, include=FALSE, message = F}
knitr::opts_chunk$set(echo = TRUE)

library(knitr)
library(dplyr)
library(ggplot2)
library(directlabels)

```

### 3.1

#### a)

\begin{align*}
Pr(Y_1 = y_1, ..., Y_100 = y_100 \mid \theta) &= \theta^{\sum Y_i}(1-\theta)^{100 - \sum Y_i}  \\ 
Pr(\sum Y_i = y \mid \theta) & = \binom{n}{y} \theta^{\sum Y_i}(1-\theta)^{100 - \sum Y_i}\\
\end{align*}



#### b)  

Given that $\sum Y_i = 57$ and $\theta \in {0.1 ... 1.0}$, $Pr(\sum Y_i \mid \theta)$ is a binomial distrbution.

```{r}
y = 57
n = 100
theta.1 = seq(0.0, 1.0, by = 0.1)
py.theta.1 = dbinom(y,n, theta.1)
table1 = matrix(c(theta.1, py.theta.1), 11, 2)
colnames(table1) = c('prior', 'y conditional on theta')
kable(table1)

df1 = as.data.frame(table1)

ggplot(df1, aes(x= theta.1, y= py.theta.1)) + 
  geom_line(aes(col = "p(y|theta)")) + 
  labs(x = "theta", y= "conditional prob of y")
```


#### c)  

$p(\theta \mid \sum Y_i = 57) = \frac{p(\sum Y_i \mid \theta)p(\theta)}{\sum_{\theta} p(\sum Y_i \mid \theta)}$

```{r}
ptheta.1 = rep(1/length(theta.1), length(theta.1))
py.theta.3 = dbinom(y,n, theta.1)
joint_y_theta = ptheta.1 * py.theta.1
poster2 = joint_y_theta / sum(joint_y_theta)
table2 = matrix(c(theta.1, poster2), 11, 2)
colnames(table2) = c('prior', 'posterior')
kable(table2)
df2 = as.data.frame(table2)

ggplot(df2, aes(x= prior, y= posterior)) + 
  geom_line(aes(col = "p(theta|y)")) + 
  geom_hline(yintercept = 1/length(theta.1), col = "blue") + 
  labs(x = "prior", y= "posterior") + 
  annotate("text", 0.05, 0.13, label = "p(theta)")
```


#### d)  


```{r}
y = 57
n = 100
theta.2 = seq(0, 1, by = 0.01)
py.theta.2 = dbinom(y,n, theta.2)
df3 <- as.data.frame(theta.2)
df3 <- cbind(df3, py.theta.2)
colnames(df3) = c('prior', 'joint')


ggplot(df3, aes(x= prior, y= joint)) + 
  geom_line(aes(col = "p(theta, y)")) +
  labs(x = "prior", y= "joint")
```

#### e)

```{r}
atheta.y.2 = 1 + y
btheta.y.2 = 1 + n - y
theta.2 = seq(0.01, 0.99, by = 0.01)
ptheta.2 = dbeta(theta.2, 1,1)
ptheta.y.2 = dbeta(theta.2, atheta.y.2, btheta.y.2)
table4 = matrix(c(theta.2, ptheta.y.2), 99, 2)
colnames(table4) = c('prior', 'posterior')

df4 = as.data.frame(table4)

ggplot(df4, aes(x= prior, y= posterior)) + 
  geom_line(aes(col = "p(theta|y)")) +
  labs(x = "prior", y= "posterior")
```


Overall speaking, the four graphs share the same shapes but different scales. with part c and e, graphs are the posterior density, but e has smaller variance.
With continues uniform as prior in part d and e, curves look more smoothier than first two graphs with discrete probability of $\frac{1}{11}$. 


### 3.2

```{r}
y <- 57 
n <- 100
theta.0 <- seq(0.1, 0.9, by = 0.1) 
n.0 <- c(1,2,8,16,32)
post.means <- matrix(NA,
                     nrow = length(theta.0),
                     ncol = length(n.0))
for (i in 1:length(theta.0)) { 
  for (j in 1:length(n.0)) {
  a.0.ij <- theta.0[i]*n.0[j]
  b.0.ij <- (1 - theta.0[i])*n.0[j]
  a.post.ij <- a.0.ij + y
  b.post.ij <- b.0.ij + n - y
  post.means[i, j] <- 1 - pbeta(0.5, a.post.ij, b.post.ij) }
}

countour.df <- as.data.frame(rep(n.0, c(rep(9,5))))
countour.df <- cbind(countour.df, rep(theta.0, 5))
countour.df <- cbind(countour.df, as.numeric(post.means))

colnames(countour.df) <- c("n.0", "theta.0", "prob")

p <-ggplot(countour.df, aes(x = theta.0, y = n.0, z = prob)) + stat_contour(aes(colour = ..level..), breaks = 0.05*(7:19)) + 
  xlim(0.1,0.9) + ggtitle("Pr( theta, > 0.5 | y = 57)") +
  theme(plot.title = element_text(hjust = 0.5))

direct.label(p,"bottom.pieces")
```

As we know, posterior probability is determined by our prior belife on $\theta_0$ as well as the strength towards that belief $n_0$.

In this case, we want to know that if $p(\theta > 0.5 \mid y = 57)$, and we can see from the graph that posterior mean is not so sensitive to our prior belief $\theta_0$; whereas, if we have strong belief(high $n_0$) on lower $\theta_0$ value, $p(\theta > 0.5 \mid y = 57)$ is likely lower than 0.5. However, people with weak belief likely lie in the range of greater than 80%. 

### Math Problem

Given $p(H \mid c_1) = 0.6, p(H \mid c_2) = 0.4$, first two are tails. 
first, we try to calculate $p(c_1 \mid TT)$ and  $p(c_2 \mid TT)$ 

\begin{align*}
p(c_1 \mid TT)  = \frac{p(c_1,TT)}{p(TT)}  &=  \frac{p(c_1,TT)}{p(c_1,TT) + p(c_2,TT)} \\
& = \frac{0.4\times 0.4}{0.4 \times 0.4 + 0.6 \times 0.6} \\
& \approx 0.3077 \\
\end{align*}

so $$p(c_2 \mid TT) = 1 - p(c_1 \mid TT) = 0.6923$$

After all, we can calculate exptectation using formula $\mu = p(c_1 \mid TT) \mu_1 + p(c_2 \mid TT) \mu_2$. In this case, we know the event is geometric distrbution, hence $$\mu_1 = \frac{1}{0.6} \approx 1.67$$ and $$\mu_2 = \frac{1}{0.4} = 2.5$$.

Therefore, $\mu = 0.3077 \times 1.67 + 0.6923 \times 2.5 = 2.245$
